---
title: "HW 3"
author: "Sam Loewen"
date: "10/7/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(p8105.datasets)
library(knitr)
data("instacart")
data("brfss_smart2010")

```
## Problem 1

The dataset `instacart` has `r nrow(instacart)` observations, with `r ncol(instacart)` variables. The data tracks products and thier corresponding order, providing information like `order_number`, `product_name`, and its corresponding `aisle`. We can learn things like the proportion of products that were reordered (`r mean(pull(instacart, reordered))`) or the average number of days since prior order (`r mean(pull(instacart, days_since_prior_order))`).


*   How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
group_by (aisle_id) %>% 
  summarize (n_obs = n()) %>% 
arrange(desc(n_obs))
```
There are 134 aisles, and aisles 83, 24, and 123 are the most ordered from.

*   Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>%
  group_by (aisle_id) %>% 
  summarize (n_obs = n()) %>% 
  filter(n_obs > 10000) %>% 
  arrange(aisle_id) %>% 
    ggplot(aes(x = aisle_id, y = n_obs, color=n_obs)) + 
    geom_point() +
    labs(title = "Number of items ordered per aisle", 
    x = "Aisle Number", 
    y = "Nmber items ordered")
```

*   Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
group_by (aisle, product_name) %>% 
  summarize (n_obs = n()) %>% 
filter (aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
top_n(3, n_obs) %>% 
knitr::kable()
```


*   Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


```{r}
instacart %>% 
group_by (product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
filter (product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) %>% 
knitr::kable()  
```


## Problem 2

First, do some data cleaning: 

*   format the data to use appropriate variable names;
*   focus on the “Overall Health” topic
*   include only responses from “Excellent” to “Poor”
*   organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
brfss_smart2010_1 =
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter (topic == "Overall Health",
          response == "Excellent" | 
          response == "Very good" | 
          response == "Good" | 
          response == "Fair" |
          response == "Poor") %>% 
  mutate(response = factor(response, labels = c("Poor", "Fair", "Good", "Very Good", "Excellent")))
```

Using this dataset, do or answer the following (commenting on the results of each):

*   In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_smart2010_1 %>% 
  group_by(year, locationabbr) %>% 
  summarize (n_loc = n_distinct(locationdesc),
             mean_dv = mean(data_value, na.rm = TRUE)) %>% 
filter(year == 2002 | year == 2010,
       n_loc >= 7) %>% 
select (-n_loc) %>% 
arrange(year, locationabbr)

```
In 2002, CT, FL, MA, NC, NJ, and PA all were observed in 7 or more locations. In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA were observed in 7 or more locations.

*   Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_smart2010_2 =
brfss_smart2010_1 %>% 
  group_by(year, locationabbr, response) %>% 
  summarize (mean_dv = mean(data_value, na.rm = TRUE)) %>% 
filter(response == "Excellent") %>% 
select (-response) %>% 
arrange(year, locationabbr)

brfss_smart2010_2

brfss_smart2010_2 %>% 
  ggplot(aes(x=year, y=mean_dv, group = locationabbr, color = locationabbr)) + geom_point() + geom_line() +
  labs(title = "Mean data value over time, by state", 
    x = "year", 
    y = "data value")
```

The mean data values across states range from `r min(pull(brfss_smart2010_2,mean_dv))` to `r max(pull(brfss_smart2010_2,mean_dv))`. The meadian values by year are: 2002: `r median(pull(filter(brfss_smart2010_2,year==2002),mean_dv))`, 2004: `r median(pull(filter(brfss_smart2010_2,year==2004),mean_dv))`, 2006: `r median(pull(filter(brfss_smart2010_2,year==2006),mean_dv))`, 2008: `r median(pull(filter(brfss_smart2010_2,year==2008),mean_dv))`, and 2010: `r median(pull(filter(brfss_smart2010_2,year==2010),mean_dv))`. 

*   Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.


```{r}
brfss_smart2010_1 %>% 
  filter(locationabbr == "NY", 
         year == 2006 | year == 2010) %>% 
ggplot(aes(x = response, y = data_value, color = response)) + 
  geom_boxplot() + facet_grid(. ~ year)  +
  labs(title = "Distribution of data values by repsonse level, NY", 
    x = "response", 
    y = "data value")
```

In 2006, the largest range of values within a category was in `Good`. The smallest distrubution was among `Very Good`, which also has the lowest data values. In 2010 the highest data values and the largest distributions of values was in the `Excellent` group. Again the smallest values and smallest distribution was in the `Very Good` group.

## Problem 3

*   Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel = 
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(day = factor(day, labels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")),
         weekend = as.numeric(if_else(day == "Saturday" | day == "Sunday", "1", "0"))) %>% 
  select (week, day_id, day, weekend, everything()) %>% 
  arrange (week, day)
```

The dataset `accel` has `r nrow(accel)` observations, with `r ncol(accel)` variables. The data tracks one man's activity every day for 5 weeks, reporting activity every minute. A unit of activity is captued every minute and is represented in our data set with varaibles `acitvity_1` through `activity_1440`.

*   Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_2 =
  accel %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    values_to = "activity") %>% 
  group_by(week, day_id, day, weekend) %>% 
  summarize(daily_tot = sum(activity))
knitr::kable(accel_2)
```

This data show the mean activity per day: Monday - `r mean(pull(filter(accel_2, day=="Monday"), daily_tot))`, Tuesday - `r mean(pull(filter(accel_2, day=="Tuesday"), daily_tot))`, Wednesday - `r mean(pull(filter(accel_2, day=="Wednesday"), daily_tot))`, Thursday - `r mean(pull(filter(accel_2, day=="Thursday"), daily_tot))`, Friday - `r mean(pull(filter(accel_2, day=="Friday"), daily_tot))`, Saturday - `r mean(pull(filter(accel_2, day=="Saturday"), daily_tot))`, and Sunday - `r mean(pull(filter(accel_2, day=="Sunday"), daily_tot))`. This indicates that Mondays are the most active day of the week.

The mean weekday value is `r mean(pull(filter(accel_2, weekend==0), daily_tot))` and mean weekend value is `r mean(pull(filter(accel_2, weekend==1), daily_tot))`, showing that weekends are slightly more active than weekdays.

*   Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_plot = 
  accel %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    values_to = "activity") %>% 
ggplot(aes(x=day_id, y=activity, group = minute, color = day)) + 
  geom_point() + geom_line() +
  labs(title = "24 hour activity over 5 weeks", 
    x = "day", 
    y = "activity unit")
accel_plot
```

 This chart appears to show that activity spikes in the middle of the day, as is to be expected given typical sleep schedules. This person appears to be slightly more active over the weekends, but they appeared to have one very active weekend that may be skewing the results. Similarly this person had one very inactive Friday-Saturday that seems like an outlier among their typical activity pattern.
